\documentclass[12pt, oneside]{amsart}
\usepackage{amsmath,amsfonts, amssymb, xcolor}
\usepackage{fullpage}
\newcommand{\Z}{\mathbb Z}
\newcommand{\R}{\mathbb R}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\spn}{span}

\usepackage[most]{tcolorbox} % powerful colored boxes
\usepackage{xcolor}          % colors
\newtcolorbox{solution}{
  colframe=green!50!black,   % medium green border
  coltext=black,             % body text in black for readability
  boxrule=0.4pt,             % thin, professional border
  enhanced,                  % enable advanced drawing features
  left=6pt, right=6pt, top=6pt, bottom=6pt % padding
}


\theoremstyle{definition}
\newtheorem{prob}{Problem}

\title{Math 223 Fall 2025 - Homework 6}
\author{due October 28 at 11:59pm}
\pagenumbering{gobble}
\begin{document}
\maketitle
Each problem is worth 10 points.

\begin{prob}
    The \emph{transpose} of an $m\times n$ matrix $A$ ($m$ rows, $n$ columns) is the $n\times m$ matrix $A^T$ ($n$ rows, $m$ columns) such that the columns of $A^T$ are the rows of $A$ (and so the rows of $A^T$ are the columns of $A$).

    \begin{enumerate}
        \item Compute the transpose of $A = \left[\begin{matrix}
        4 & 0 & 2 & 2 \\
        0 & 1 & 1 & 2
        \end{matrix}\right]$.
        
        \begin{solution}
        When we compute the transpose of a matrix, we swap the rows and columns. The columns of $A^T$ are the rows of $A$.
        
        \[
        A^T = \begin{bmatrix}
        4 & 0 \\
        0 & 1 \\
        2 & 1 \\
        2 & 2
        \end{bmatrix}
        \]
        \end{solution}
    \end{enumerate}
    Let $A$ be an $m\times n$ matrix and $B$ an $n\times k$ matrix.
    \begin{enumerate}
    \setcounter{enumi}{2}
    \item Check that matrices $(AB)^T$ and $B^T A^T$ have the same size.
    
    \begin{solution}
    
    First, we determine the size of $(AB)^T$:
    
    Since $A$ is $m \times n$ and $B$ is $n \times k$, the product $AB$ is an $m \times k$ matrix.
    
    When we transpose $AB$, the rows and columns are swapped. Therefore $(AB)^T$ is a $k \times m$ matrix.
    
    \vspace{0.3cm}
    
    Next, we determine the size of $B^T A^T$:
    
    Since $B$ is $n \times k$, its transpose $B^T$ is a $k \times n$ matrix.
    
    Since $A$ is $m \times n$, its transpose $A^T$ is an $n \times m$ matrix.
    
    The product $B^T A^T$ multiplies a $k \times n$ matrix by an $n \times m$ matrix resulting in a $k \times m$ matrix.
    
    \vspace{0.3cm}
    
    Finally, both $(AB)^T$ and $B^T A^T$ are $k \times m$ matrices, so they have the same size.
    \end{solution}
    
    \item Show that $(AB)^T = B^T A^T$. 
    (Describe how the $(i,j)$-entry of $(AB)^T$ is obtained from rows/columns of $A$ and $B$, and similarly describe how the $(i,j)$-entry of $B^T A^T$ is obtained, and compare these.)
    
    \begin{solution}
    We show that $(AB)^T = B^T A^T$ by comparing their $(i,j)$-entries $(AB)^T]_{i,j}$ and $[B^T A^T]_{i,j}$.
    
    \vspace{0.3cm}
    
    By definition of transpose:
    
    \[
    (AB)^T_{j,i} = (A B)_{i,j}
    \]
    
    The $(j,i)$-entry of $(AB)^T$ is computed by taking row $j$ of $A$ and dotting it with column $i$ of $B$:
    \[
    [(AB)^T]_{i,j} = [AB]_{j,i} = \sum_{\ell=1}^{n} A_{j,\ell} \cdot B_{\ell,i}
    \]
    
    \vspace{0.3cm}
    
    We compute the $(i,j)$-entry of $B^T A^T$ by taking row $i$ of $B^T$ and dotting it with column $j$ of $A^T$. 
    
    \hspace{2em} Row $i$ of $B^T$ = Column $i$ of $B$ = $(B_{1,i}, B_{2,i}, \ldots, B_{n,i})$

    \hspace{2em} Column $j$ of $A^T$ = Row $j$ of $A$ = $(A_{j,1}, A_{j,2}, \ldots, A_{j,n})$

    
    Therefore:
    $[B^T A^T]_{i,j} = \sum_{\ell=1}^{n} B_{\ell,i} \cdot A_{j,\ell}$
    
    \vspace{0.3cm}

    Since $A_{j,\ell}$ and $B_{\ell,i}$ are single elements in their respective matrices and thus are scalars, we know multiplication of these scalars is commutative. Thus:
    \[
      A_{j,\ell} \cdot B_{\ell,i} = B_{\ell,i} \cdot A_{j,\ell}
    \]
    \[
    \text{for each: } \ell
    \]
    
    Thus:
    \[
    [(AB)^T]_{i,j} = \sum_{\ell=1}^{n} A_{j,\ell} \cdot B_{\ell,i} = \sum_{\ell=1}^{n} B_{\ell,i} \cdot A_{j,\ell} = [B^T A^T]_{i,j}
    \]
    
    
    Because the $(i,j)$-entries are equal for all $i$ and $j$, we conclude that $(AB)^T = B^T A^T$. \qed
    \end{solution}

    \end{enumerate}

\end{prob}


\begin{prob} Let $A$ be an $n\times n$ matrix. 
    \begin{enumerate}
    \item Let $A = E_1\cdots E_\ell$ be the product of elementary matrices. Deduce from Problem 1 that $A^T = E_\ell^T \cdots E_1^T$.
    
    \begin{solution}
    
    
    We write our matrix product as: 
    
    \[
    A = E_1 E_2 \cdots E_\ell = (E_1 E_2 \cdots E_{\ell-1}) E_\ell
    \]

    Taking the transpose of both sides:
    
    \[
      A^T = (E_1 E_2 \cdots E_\ell)^T = ((E_1 E_2 \cdots E_{\ell-1}) E_\ell)^T
    \]
    
    Now we can apply the result of problem 1, $(AB)^T = B^T A^T$, with the product $(E_1 \cdots E_{\ell-1})$ as the first matrix and $E_\ell$ as the second matrix:
    
    
    \[
      A^T = E_\ell^T (E_1 \cdots E_{\ell-1})^T
    \]
    
    Repeatedly applying this step, we remove the rightmost $E_i$ from inside the parenthesis and places its transpose on the left. 

    After applying this to each element, $\ell-1$ times, we get the result:

    \[
      A^T = E_\ell^T E_{\ell-1}^T \cdots E_2^T E_1^T
    \]
    
    For example with $\ell = 3$:
    \[
      A = E_1 E_2 E_3 \implies A^T = ((E_1 E_2) E_3)^T = E_3^T (E_1 E_2)^T = E_3^T E_2^T E_1^T
    \]
    \end{solution}
    
    \item Show that for every elementary matrix $E$, $E^T$ is also an elementary matrix, and $\det E = \det E^T$.
    
    \begin{solution}
    
    Elementary matrices come from doing one of three row operations on the identity matrix. We will show that $E^T$ is an elementary matrix and $\det E = \det E^T$ for each of these three cases.
    
    \vspace{0.3cm}
    
    \textbf{Row swap operation}
    
    Consider the elementary matrix, $E$ where rows $i$ and $j$ are swapped from the identity matrix.
    
    Because the matrix transpose operation swaps the rows and columns of a matrix, when we transpose $E$ the operation of swapping rows $i$ and $j$ which we performed on the identity matrix, becomes swapping columns $i$ and $j$.
    
    Since the identity matrix is symmetric, swapping columns $i$ and $j$ of $I$ produces the same matrix as swapping rows $i$ and $j$ of $I$. Thus:
    \[
      E^T = E
    \]

    As these two matrices are equal, their determinants are equal:
    \[
    \det E = \det E^T
    \]
    
    \vspace{0.3cm}
    
    \textbf{Row scaling operation}
    
    Consider the elementary matrix, $E$, obtained by scaling row $i$ by $c \in F$ in the identity matrix.
    
    The matrix $E$ is diagonal as it is equal to the identity matrix with a single entry scaled. 

    Thus its entries along the diagonal are: $1, 1, \ldots, c, \ldots, 1$ where $c$ is in position $(i,i)$.
    
    Since $E$ is diagonal, transposing it produces the same matrix:
    \[
      E^T = E
    \]
    
    As these two matrices are equal, their determinants are equal:
    \[
    \det E = \det E^T
    \]
    
    \vspace{0.3cm}
    
    \textbf{Row addition operation}
    
    Consider the elementary matrix, $E$ obtained by adding $c$ times row $i$ to row $j$ in the identity matrix.
    
    The matrix $E$ has 1's on the diagonal, a $c$ in position $(j, i)$, and 0's elsewhere.
    
    When we transpose, the $c$ moves from position $(j, i)$ to position $(i, j)$. Therefore $E^T$ has 1's on the diagonal and a $c$ in position $(i, j)$.
    
    This means $E^T$ is equivelant to the identity matrix with the row operation of adding $c$ times row $j$ to row $i$. As that is a single row addition operation performed on the identity matrix, $E^T$ is an elementary matrix.
    
    We also know that row addition operations do not change the determinant, so:
    \[
    \det E = 1 = \det E^T
    \]
    
    \vspace{0.3cm}
    
    We have shown that in all three cases, $E^T$ is an elementary matrix and $\det E = \det E^T$. 
    \end{solution}
    
    \item Prove that for every $n\times n$ matrix $A$ we have $\det A = \det A^T$.
    
    \begin{solution}
    We prove this by splitting it into two cases: when $A$ is invertible and when $A$ is not invertible.
    
    \vspace{0.3cm}
    
    \textbf{$A$ is invertible}
    
    If $A$ is invertible, then $A$ can be written as a product of elementary matrices:
    \[
    A = E_1 E_2 \cdots E_\ell
    \]
    
    We can take the determinant of A, and because determinants can be multiplied linearly, we can split them into determinants of each elementary matrix:
    \[
    \det A = \det(E_1 E_2 \cdots E_\ell) = \det(E_1) \cdot \det(E_2) \cdots \det(E_\ell)
    \]
    
    We can do the same split over the transpose of $A$. From part 1 of this problem we have already shown that ;
    
    \[A^T = E_\ell^T E_{\ell-1}^T \cdots E_2^T E_1^T\]
    
    Now we apply the same splitting of the determinants over multiplication:
    \[
    \det A^T = \det(E_\ell^T E_{\ell-1}^T \cdots E_1^T) = \det(E_\ell^T) \cdot \det(E_{\ell-1}^T) \cdots \det(E_1^T)
    \]
    
    From part 2 of this problem, we know $\det(E_i^T) = \det(E_i)$ for any elementary matrix $E_i$. Thus we can substitute:
    \[
    \det A^T = \det(E_\ell^T) \cdot \det(E_{\ell-1}^T) \cdots \det(E_1^T) = \det(E_\ell) \cdot \det(E_{\ell-1}) \cdots \det(E_1)
    \]
    
    Finally, as the determinants are just scalars, we can rearrange our product of $\det(A^T)$ to the same form as $\det(A)$:
    \[
    \det A^T = \det(E_1) \cdot \det(E_2) \cdots \det(E_\ell) = \det A
    \]
    
    \vspace{0.3cm}
    
    \textbf{$A$ is not invertible}
    
    If $A$ is not invertible, then $\det A = 0$.
    This implies that the rows or columns of $A$ are linearly dependent. 
    
    Thus when we tranpose $A$, the rows and columns swap, meaning that if the rows were previously linearly dependent, the columns are now linearly dependent (and vice versa).

    In either situation, as the rows or columns of $A^T$ are linearly dependent, then $A^T$ is not invertible and we know $\det A^T = 0$.
    Thus: 

    \[
    \det A^T = 0 = \det A
    \]
    
    \vspace{0.3cm}
    
    \end{solution}
    \end{enumerate}
\end{prob}

\begin{prob} True or False. Justify your answers.
\begin{enumerate}
    \item  $\det (A+B) = \det A  +\det B$ where $A,B$ are $n\times n$ matrices.
    
    \begin{solution}
    False, which can be proved with the counter example:
    
    \[
    A = \begin{bmatrix} 1 & 0 \\ 0 & 0 \end{bmatrix}, \quad B = \begin{bmatrix} 0 & 0 \\ 0 & 1 \end{bmatrix}
    \]
    
    Finding the determinants of $A$ and $B$:
    \[
    \det A = 0 \quad \det B = 0
    \]
    \[
    \det A + \det B = 0 + 0 = 0
    \]
    
    However, when we add the matrices:
    \[
    A + B = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
    \]
    
    \[
    \det(A+B) = 1 
    \]

    Finally:
  \[
  \det(A+B) \neq 0 = \det A + \det B
  \]

    \end{solution}
    
    \item $\det c\cdot A = c^n \det A$ where $A$ is an $n\times n$ matrix, $c\in F$ and $c\cdot A$ is obtained from $A$ by multiplying each entry by $c$.
    
    \begin{solution}
    We know from the properties of determinants that multiplying a single row by a scalar $c$ multiplies the determinant by $c$.
    
    When we compute $c \cdot A$, we are multiplying every entry of $A$ by $c$, which is the same as multiplying each of the $n$ rows by $c$.
    
    Since each time we multiply a row by $c$, this multiplies the determinant by $c$, and we perform this operation $n$ times (once for each row), the determinant is multiplied by $c$ a total of $n$ times.

    Thus:
    \[
    \det(c \cdot A) = c^n \det A
    \]
    
    Therefore the statement is true.
    \end{solution}
\end{enumerate}
\end{prob}

\begin{prob} Compute the determinant.
\begin{enumerate}
    \item $A_1= \left[\begin{matrix}
        1 & 0 & -2 & 3 \\
        -3 & 1 & 1 & 2 \\
        0 & 4 & -1 & 1 \\
        2 & 3 & 0 & 1
    \end{matrix}\right]$\\\\
    
    \begin{solution}
    We compute the determinant via row reduction to an upper triangular matrix and then take the product of the diagonal:
    
    Row Operations:
    
    $R_2 = R_2 + 3R_1$
    
    $R_4 = R_4 - 2R_1$
    
    $R_3 = R_3 - 4R_2$
    
    $R_4 = R_4 - 3R_2$
    
    $R_4 = R_4 - R_3$

    This results in the upper triangular matrix:
    
    \[
    \begin{bmatrix}
    1 & 0 & -2 & 3 \\
    0 & 1 & -5 & 11 \\
    0 & 0 & 19 & -43 \\
    0 & 0 & 0 & 5
    \end{bmatrix}
    \]
    
    The determinant is the product of the diagonal entries:
    \[
    \det A_1 = 1 \cdot 1 \cdot 19 \cdot 5 = 95
    \]
    \end{solution}
    
    \item $A_2= \left[\begin{matrix}
        -1 & 3 & 2 \\4 & -8 & 1 \\
        2 & 2 & 5
    \end{matrix}\right]$\\\\
    
    \begin{solution}
    We use the known formula for a cofactor expansion for a 3×3 matrix, which gives the determinant:
    
    \[
    \det A_2 = -1 \begin{vmatrix} -8 & 1 \\ 2 & 5 \end{vmatrix} - 3 \begin{vmatrix} 4 & 1 \\ 2 & 5 \end{vmatrix} + 2 \begin{vmatrix} 4 & -8 \\ 2 & 2 \end{vmatrix}
    \]
    
    \[
    = -1[(-8)(5) - (1)(2)] - 3[(4)(5) - (1)(2)] + 2[(4)(2) - (-8)(2)]
    \]
    
    \[
    = -1[-42] - 3[18] + 2[24] = 42 - 54 + 48 = 36
    \]
    
    Therefore $\det A_2 = 36$.
    \end{solution}

    \item $A_3= \left[\begin{matrix}
        0 & 0 & 0 & 0 & 0 & -1 \\
        0 & 0 & 0 & 0 & 2 & 0 \\ 
        0 & 0 & 0 & 1 & 0 & 0 \\
        0 & 0 & 3 & 0 & 0 & 0 \\
        0 & 4 & 0 & 0 & 0 & 0 \\
        1 & 0 & 0 & 0 & 0 & 0 
    \end{matrix}\right]$
    
    \begin{solution}
    We compute the determinant via row reduction to a diagonal matrix and then take the product of the diagonal:
    
    Row Operations:
    
    Swap: $R_1$ and $R_6$

    Swap: $R_2$ and $R_5$

    Swap: $R_3$ and $R_4$

    This results in the upper triangular matrix:
    \[
    \begin{bmatrix}
    1 & 0 & 0 & 0 & 0 & 0 \\
    0 & 4 & 0 & 0 & 0 & 0 \\
    0 & 0 & 3 & 0 & 0 & 0 \\
    0 & 0 & 0 & 1 & 0 & 0 \\
    0 & 0 & 0 & 0 & 2 & 0 \\
    0 & 0 & 0 & 0 & 0 & -1
    \end{bmatrix}
    \]
    
    The product of the diagonal entries is:
    \[
    1 \cdot 4 \cdot 3 \cdot 1 \cdot 2 \cdot (-1) = -24
    \]
    
    We performed 3 row swaps, and each swap multiplies the determinant by $-1$:
    \[
    \det A_3 = (-1)^3 \cdot (-24) = 24
    \]
    \end{solution}
\end{enumerate}
\end{prob}
\noindent \textbf{Generative AI Acknowledgment}

\noindent Generative AI was used in this assignment for help with latex syntax. This includes defining the "solution box" that wraps answers. This also includes transcribing verbatum from handwritten on-paper solutions to first-draft latex code.
\end{document}

\end{document}